---
title: '*LDlinkR*: An R Package for Rapidly Calculating Linkage Disequilibrium Statistics in Diverse Populations'
author: "Timothy A. Myers, Stephen J. Chanock and Mitchell J. Machiela"
date: "28 January 2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{*LDlinkR*: An R Package for Rapidly Calculating Linkage Disequilibrium Statistics in Diverse Populations}
  %\VignetteEngine{knitr::rmarkdown}
  \usagepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("LDlinkR")
```

***

## Description

[LDlink](https://ldlink.nci.nih.gov) is an interactive and powerful suite of web-based tools for querying germline variants in human population groups of interest to generate interactive tables and plots.  All population genotype data originates from Phase 3 (Version 5) of the [1000 Genomes Project](https://www.internationalgenome.org/) and variant RS numbers are indexed based on dbSNP 151.

*LDlinkR* is an R package developed to query and download results generated by LDlink web-based applications from the R console.  *LDlinkR* accelerates genomic research by providing efficient and user-friendly functions to programmatically interrogate pairwise linkage disequilibrium from large lists of genetic variants. 

## Installation
- The release version of *LDlinkR* can be installed from CRAN:
```{r eval=FALSE}
install.packages("LDlinkR")
```
- The development version of the *LDlinkR* package can be installed from the GitHub repository by using the *remotes* package:
```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("CBIIT/LDlinkR")
```
*LDlinkR* depends on the following packages:

- *utils*, version 3.4.2 or later
- *httr*, version 1.4.0 or later

Following installation, attach the *LDlinkR* package with:
```{r eval=FALSE}
library(LDlinkR)
```


## Personal Access Token - Required

In order to access the LDlink API via *LDlinkR*, we use a personal access token. This is a common convention followed by many APIs and emulates the more familiar HTTPS username/password or SSH keys.

You will need to:

- Make a one-time request for your personal access token from a web browser at <https://ldlink.nci.nih.gov/?tab=apiaccess>.
- Once registered, your personal access token will be emailed to you. It is a string of 12 random letters and numbers. 
- Provide your token as an argument when using *LDlinkR*.  See example below:

```{r, eval=FALSE}
LDhap(snps = c("rs3", "rs4", "rs148890987"), 
      pop = "YRI", 
      token = "YourTokenHere123")
```

**Optional:**
<br>
However, the best security practice is to store your personal access token as an environment variable where *LDlinkR* can find it and use it on your behalf but where it will not be accidentally shared with the public.  **Note:**  Modifying R startup files (such as the `.Renviron`) is for the advanced R user only.  Modification of these files in the wrong way could cause problems.  Please proceed cautiously.  Step-by-step instructions follow:

After retrieving your personal access token from your email, put your token in your `.Renviron` file. `.Renviron` is a hidden file that lives in your home directory.  The easiest way to both find and edit the `.Renviron` file is with a function from the *usethis* package.  From the R console, do:

```{r eval=FALSE}
usethis::edit_r_environ()
```
Your `.Renviron` file should open in your editor.  Add a line that looks like this:
```{r eval=FALSE}
LDLINK_TOKEN=YourTokenHere123
```
**Important**, ensure you put a line break at the end by hitting the *enter/return* key.

Save and close the `.Renviron` file.  Restart R, as environment variables are only loaded from `.Renviron` at the start of a new R session. Now, check to see that your token is available by entering:
```{r eval = FALSE}
Sys.getenv("LDLINK_TOKEN")
```

```{}
## [1] "YourTokenHere123"
```

You should see your personal access token print to the screen, as shown above. Now, *LDlinkR* function calls that use
```{r eval=FALSE}
Sys.getenv("LDLINK_TOKEN")
```

for the `token` argument in *LDlinkR* function calls will use your personal access token in a private and secure way.  This method will be used in the extended examples that follow.

***

## Functions and Examples

### LDexpress

#### Function 
```{r eval=FALSE}
LDexpress(snps, 
          pop = "CEU", 
          tissue = "ALL", 
          r2d = "r2", 
          r2d_threshold = 0.1, 
          p_threshold = 0.1, 
          win_size = 500000, 
          token = NULL, 
          file = FALSE
         )
```
Search if a list of genomic variants (or variants in LD with those variants) is associated with gene expression in tissues of interest. Quantitative trait loci data is downloaded from the [GTEx Portal](https://gtexportal.org/home/).

#### Arguments
- `snps`, between 1 - 10 variants, using an rsID or chromosome coordinate (e.g. "chr7:24966446")
- `pop`, a 1000 Genomes Project population, (e.g. YRI or CEU), multiple allowed, default = "CEU".  See the *list_pop* function in the utilities section below for available human populations and their abbreviation codes.
- `tissue`, select from 1 - 54 non-diseased tissue sites collected for the GTEx project, multiple allowed. Acceptable user input is taken either from "tissue_name_ldexpress" or "tissue_abbrev_ldexpress" (tissue abbreviation) code listed in available GTEx tissue sites using the *list_gtex_tissues* function (e.g. "ADI_SUB" for Adipose Subcutaneous). Input is case sensitive. Default = "ALL" for all available tissue types.
- `r2d`, Select either "r2" for LD R^2^ (R-squared) or "d" for LD D', default = "r2".
- `r2d_threshold`, R-squared or D' (depends on 'r2d' user input parameter) threshold for LD filtering. Any variants within -/+ of the specified genomic window and R^2^ or D' less than the threshold will be removed. Value needs to be in the range 0 to 1. Default value is 0.1.
- `p_threshold`, define the eQTL significance threshold used for returning query results. Default value is 0.1 which returns all GTEx eQTL associations with P-value less than 0.1.
- `win_size`, set genomic base pair window size for LD calculation. Specify a value greater than or equal to zero and less than or equal to 1000000 basepairs (bp). Default value is -/+ 500000 bp.
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage: Single query variant, multiple populations, multiple tissue types using tissue abbreviation

```{r}
my_output <- LDexpress(snps = "rs4",
                       pop = c("YRI", "CEU"),
                       tissue =  c("ADI_SUB", "ADI_VIS_OME"),
                       win_size = "500000",
                       token = Sys.getenv("LDLINK_TOKEN")
                      )
```

In the above example, output is a data frame stored in the variable `my_output`.
See below.

```{r}
head(my_output)
```

#### Usage: Multiple query variants, single population, a tissue type using the full LDexpress tissue name, no spaces.

```{r}
my_output <- LDexpress(snps = c("rs345", "rs456"),
                       pop = "YRI",
                       tissue =  "Adipose_Visceral_Omentum",
                       token = Sys.getenv("LDLINK_TOKEN")
  )
```

In the above example, output is a data frame stored in the variable `my_output`.
See below.

```{r}
head(my_output)
```

```{r}
tail(my_output)
```

***

### LDhap

#### Function 
```
LDhap(snps, pop="CEU", token=NULL, file = FALSE)
```
Calculates population specific haplotype frequencies of all haplotypes observed for a list of query variants. Input is a list of variant RS numbers (concatenated list) and a population group.

#### Arguments
- `snps`, a list of between 1 - 30 variants, using an rsID or chromosome coordinate (e.g. "chr7:24966446")
- `pop`, a 1000 Genomes Project population, uses three letter population code, (e.g. YRI or CEU), multiple allowed, default = "CEU"
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage: Multiple query variants, single population
```{r eval=FALSE}
LDhap(snps = c("rs3", "rs4", "rs148890987"), 
      pop = "CEU", 
      token = Sys.getenv("LDLINK_TOKEN")
     )
```

```
##   rs148890987 rs3 rs4 Count Frequency
## 1           C   C   A   176    0.8889
## 2           T   T   G    11    0.0556
## 3           T   C   A     7    0.0354
## 4           C   T   G     4    0.0202
```

<br>

#### Usage: Multiple query variants, multiple populations
```{r eval=FALSE}
LDhap(snps = c("rs3", "rs4", "rs148890987"),
      pop = c("YRI", "CEU"),
      token = Sys.getenv("LDLINK_TOKEN")
     )
```

```
##   rs148890987 rs3 rs4 Count Frequency
## 1           C   C   A   355    0.8575
## 2           C   T   G    41     0.099
## 3           T   T   G    11    0.0266
## 4           T   C   A     7    0.0169
```

Output is a table of alleles, haplotype count and haplotype frequencies. 

***

### LDmatrix

#### Function
```
LDmatrix(snps, pop = "CEU", r2d = "r2", token = NULL, file = FALSE)
```
Generates a data frame of pairwise linkage disequilibrium statistics. Input is a list of between 2 to 1000 variants. Desired output can be based on estimates of R^2^ or D'.

#### Arguments
- `snps`, list of between 2 - 1,000 variants, using an rsID or chromosome coordinate (GRCh37/hg19) (e.g. "chr7:24966446")
- `pop`, a 1000 Genomes Project population, uses three letter population code, (e.g. YRI or CEU), multiple allowed, default = "CEU"
- `r2d`, use either "r2" for pairwise R^2^ statistics or "d" for pairwise D' statistics
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage: Multiple query variants, single population, R^2^
```{r eval=FALSE}
LDmatrix(snps = c("rs496202", "rs11147477", "rs201578600"), 
         pop = "YRI", r2d = "r2", 
         token = Sys.getenv("LDLINK_TOKEN")
        )
```

```
##     RS_number rs496202 rs201578600 rs11147477
## 1    rs496202    1.000       0.660      0.504
## 2 rs201578600    0.660       1.000      0.786
## 3  rs11147477    0.504       0.786      1.000
```

<br>

#### Usage: Multiple query variants (rsID & genomic coordinates), multiple populations, D'
```{r eval=FALSE}
LDmatrix(snps = c("chr13:32444611", "rs11147477", "rs201578600"), 
         pop = c("YRI", "CEU"), r2d = "d", 
         token = Sys.getenv("LDLINK_TOKEN")
        )
```

```
##     RS_number rs496202 rs201578600 rs11147477
## 1    rs496202    1.000       0.973      0.738
## 2 rs201578600    0.973       1.000      0.971
## 3  rs11147477    0.738       0.971      1.000
```

<br>

#### Usage: Multiple query variants read from text file, multiple populations, D'
```{r}
my_variants <- read.table("variant_list.txt")
my_variants
```
Then, call *LDmatrix* with:
```{r eval=FALSE}
LDmatrix(snps = my_variants[,1], 
         pop = c("YRI", "CEU"), r2d = "d", 
         token = Sys.getenv("LDLINK_TOKEN")
        )
```

```
##    RS_number rs60676332 rs7805287 rs127 rs456 rs10239961 rs114
## 1 rs60676332      1.000     0.094 0.180 0.151      0.363 0.148
## 2  rs7805287      0.094     1.000 0.818 0.789      0.464 0.710
## 3      rs127      0.180     0.818 1.000 0.929      0.912 0.886
## 4      rs456      0.151     0.789 0.929 1.000      1.000 0.963
## 5 rs10239961      0.363     0.464 0.912 1.000      1.000 0.459
## 6      rs114      0.148     0.710 0.886 0.963      0.459 1.000
```

Output is a table with rows and columns equal to the number of query variants and pairwise linkage disequilibrium statistics.

***

### LDpair

#### Function
```
LDpair(var1, var2, pop = "CEU", token = NULL, output = "table", file = FALSE)
```
Investigates potentially correlated alleles for a pair of variants.  Input is two query variants and a 1000 Genomes Project reference population(s) of interest.

#### Arguments
- `var1`, the first RS number (rsID) or genomic coordinate (GRCh37/hg19) (e.g. "chr7:24966446"), must match a bi-allelic variant
- `var2`, the second RS number or genomic coordinate, as above, must match a bi-allelic variant
- `pop`, a 1000 Genomes Project reference population, uses three letter population code, (e.g. YRI or CEU), multiple allowed, default = "CEU"
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `output`, two output format options are available, "text", which displays a two-by-two matrix displaying haplotype counts and allele frequencies along with other statistics, or "table", which displays the same data in rows and columns, default = "table"
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage: With `output` argument set to "text"
```{r eval=FALSE}
LDpair(var1 = "rs496202", 
       var2 = "rs11147477", 
       pop = "YRI", 
       token = Sys.getenv("LDLINK_TOKEN"), 
       output = "text"
      )
```

```
## Query SNPs:
## rs496202 (chr13:32444611)
## rs11147477 (chr13:32509120)
## 
## YRI Haplotypes:
##                rs11147477
##                C       T
##              -----------------
##            C | 11    | 26    | 37    (0.171)
## rs496202     -----------------
##            G | 173   | 6     | 179   (0.829)
##              -----------------
##                184     32      216
##               (0.852) (0.148)
## 
##           G_C: 173 (0.801)
##           C_T: 26 (0.12)
##           C_C: 11 (0.051)
##           G_T: 6 (0.028)
## 
##           D': 0.7737
##           R2: 0.5037
##       Chi-sq: 108.8005
##      p-value: <0.0001
## 
## rs496202(C) allele is correlated with rs11147477(T) allele
## rs496202(G) allele is correlated with rs11147477(C) allele
```

<br>

#### Usage: With no `output` argument option specified, using default "table"
```{r eval=FALSE}
LDpair(var1 = "rs496202", 
       var2 = "rs11147477", 
       pop = "YRI", 
       token = Sys.getenv("LDLINK_TOKEN")
      )
```

```
##       var1       var2 pops       var1_pos       var2_pos var1_a1 var1_a2
## 1 rs496202 rs11147477  YRI chr13:32444611 chr13:32509120       C       G
##   var1_a1_freq var1_a2_freq var2_a1 var2_a2 var2_a1_freq var2_a2_freq
## 1        0.171        0.829       C       T        0.852        0.148
##   d_prime     r2    chisq p_val
## 1  0.7737 0.5037 108.8005 1e-04
##                                           corr_alleles
## 1 rs496202(C)-rs11147477(T), rs496202(G)-rs11147477(C)
```

Output of the `output` argument "text" option is a two-by-two contingency table displaying haplotype counts and allele frequencies of the two query variants. Also displayed are calculated metrics of linkage disequilibrium including: D prime (D'), R square (R^2^), and goodness-of-fit (Chi-square and p-value). Goodness-of-fit tests for deviations of expected haplotype frequencies based on allele frequencies. Correlated alleles are reported if linkage disequilibrium is present (R^2^ > 0.1). If linkage equilibrium, no alleles are reported. 

Output from the `output` argument "table" option converts the data from the two-by-two contingency table into a data frame.

***

### LDpop

#### Function
```
LDpop(var1, var2, pop = "CEU", r2d = "r2", token = NULL, file = FALSE)
```
Investigates allele frequencies and linkage disequilibrium patterns across 1000G populations.

#### Arguments
- `var1`, the first RS number (rsID) or genomic coordinate (GRCh37/hg19) (e.g. "chr7:24966446"), must match a bi-allelic variant
- `var2`, the second RS number or genomic coordinate, as above, must match a bi-allelic variant
- `pop`, a 1000 Genomes Project reference population, uses three letter population code, (e.g. YRI or CEU), multiple allowed, default = "CEU"
- `r2d`, use "r2" if desired output is based on estimated R^2^ or "d" if D'
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage
```{r eval=FALSE}
LDpop(var1 = "rs496202", 
      var2 = "rs11147477", 
      pop = "YRI", 
      r2d = "r2", 
      token = Sys.getenv("LDLINK_TOKEN")
     )
```

```
##   Population   N rs496202_Allele_Freq rs11147477_Allele_Freq     R2     D'
## 1        YRI 108 G: 82.87%, C: 17.13%   C: 85.19%, T: 14.81% 0.5037 0.7737
```

***

### LDproxy

#### Function
```
LDproxy(snp, pop = "CEU", r2d = "r2", token = NULL, file = FALSE)
```
Explore proxy and putative functional variants for a single query variant. Input is a single RS number and a population group.  Depending on the number of query populations, this function could take some time to run.

#### Arguments
- `snp`, an RS number (rsID) or chromosome coordinate (GRCh37/hg19) (e.g. "chr7:24966446"), one per query, RS number must match a bi-allelic variant 
- `pop`, a 1000 Genomes Project reference population, uses three letter population code, (e.g. YRI or CEU), multiple allowed, default = "CEU"
- `r2d`, use "r2" if desired output is based on estimated R^2^ or "d" if D'
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage: single reference population
```{r eval=FALSE}
my_proxies <- LDproxy(snp = "rs456", 
                      pop = "YRI", 
                      r2d = "r2", 
                      token = Sys.getenv("LDLINK_TOKEN")
                     )
```

Output is a data frame stored in the variable `my_proxies` with 2455 rows and 10 columns with data. 
```{r eval=FALSE}
head(my_proxies)
```

```
##    RS_Number         Coord Alleles    MAF Distance Dprime     R2
## 1      rs456 chr7:24962419   (G/C) 0.1944        0      1 1.0000
## 2      rs457 chr7:24962426   (T/C) 0.1944        7      1 1.0000
## 3 rs28475742 chr7:24964633   (G/T) 0.1944     2214      1 1.0000
## 4      rs123 chr7:24966446   (C/A) 0.1944     4027      1 1.0000
## 5      rs125 chr7:24959703   (C/T) 0.2037    -2716      1 0.9436
## 6      rs128 chr7:24958977   (C/T) 0.2037    -3442      1 0.9436
##   Correlated_Alleles RegulomeDB Function
## 1            G=G,C=C          5     <NA>
## 2            G=T,C=C          5     <NA>
## 3            G=G,C=T          4     <NA>
## 4            G=C,C=A         1f     <NA>
## 5            G=C,C=T          5     <NA>
## 6            G=C,C=T          7     <NA>
```

Includes information on all variants -/+ 500 Kb of the query variant with a pairwise R^2^ value greater than 0.01.

***

### LDproxy_batch

#### Function
```
LDproxy_batch(snp, pop = "CEU", r2d = "r2", token = NULL, append = FALSE)
```
Query *LDproxy* using a list of query variants. *LDproxy_batch* will make sequential queries, one query per variant.  Concurrent queries are not permitted by the LDlink API.  Output is saved as text file(s) to the current working directory. Depending on the number of query variants and reference populations selected, this function could time some time to run.

#### Arguments
- `snp`, a character string or data frame listing RS numbers (rsID) or chromosome coordinates (GRCh37/hg19) (e.g. "chr7:24966446"), one per line.
- `pop`, a 1000 Genomes Project reference population, uses three letter population code, (e.g. YRI or CEU), multiple allowed, default = "CEU"
- `r2d`, use "r2" if desired output is based on estimated R^2^ or "d" if D'
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `append`, a logical, if TRUE, output for each query variant is appended to a single text file and saved to the current working directory.  If FALSE, output for each query variant is saved in its own text file with the query variant as the filename.  Default value is FALSE.

#### Usage: multiple variants, default `pop` and `r2d` 
The list of query variants passed to *LDproxy_batch* can be stored as a character string.
```{r eval=FALSE}
LDproxy_batch(snp = c("rs456", "rs114", "rs127"), 
              token = Sys.getenv("LDLINK_TOKEN")
             )
```
Or, a longer list of variants can be read into a data frame from a text file and passed into *LDproxy_batch*. The list should be in a simple text file, one query variant per line.  For example:
```{r}
my_variants <- read.table("variant_list.txt")
my_variants
```
Then, call *LDproxy_batch* with:
```{r eval=FALSE}
LDproxy_batch(snp = my_variants, 
              token = Sys.getenv("LDLINK_TOKEN")
             )
```
Output not displayed.  All output from *LDproxy_batch* is saved to a text file(s) in the current working directory.

***

### LDtrait

#### Function 
```{r eval=FALSE}
LDtrait(snps,
        pop = "CEU",
        r2d = "r2",
        r2d_threshold = 0.1,
        win_size = 500000,
        token = NULL,
        file = FALSE
        )
```
Search if a list of variants (or variants in LD with those variants) have been previously associated with a trait or disease. Trait and disease data is updated nightly from the [GWAS Catalog](https://www.ebi.ac.uk/gwas/docs/file-downloads).

#### Arguments
- `snps`, between 1 - 50 variants, using an rsID or chromosome coordinate (GRCh37)(e.g. "chr7:24966446"). All input variants must match a bi-allelic variant.
- `pop`, a [1000 Genomes Project](https://www.internationalgenome.org/) population, (e.g. YRI or CEU), multiple allowed, default = "CEU".  See the *list_pop* function in the utilities section below for available human populations and their abbreviation codes.
- `r2d`, use "r2" to filter desired output from a threshold based on estimated LD R2 (R squared) or "d" for LD D' (D-prime), default = "r2".
- `r2d_threshold`, R-squared or D' (depends on 'r2d' user input parameter) threshold for LD filtering. Any variants within -/+ of the specified genomic window and R^2^ or D' less than the threshold will be removed. Value needs to be in the range 0 to 1. Default value is 0.1.
- `win_size`, set genomic base pair window size for LD calculation. Specify a value greater than or equal to zero and less than or equal to 1000000 basepairs (bp). Default value is -/+ 500000 bp.
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage: Single query variant and multiple reference populations 

```{r}
LDtrait(snps = "rs456",
        pop = c("YRI", "CEU"),
        token = Sys.getenv("LDLINK_TOKEN")
)
```

#### Usage:  Multiple query variants, multiple reference populations and *win_size* set to 750000 basepairs (bp).

```{r}
LDtrait(snps = c("rs114", "rs496202", "rs345"),
        pop = c("YRI", "CHB", "CEU"),
        win_size = "750000",
        token = Sys.getenv("LDLINK_TOKEN")
)
```

***

### SNPchip

#### Function
```
SNPchip(snps, chip = "ALL", token = NULL, file = FALSE)
```
Used to find commercial genotyping chip arrays for variants.  Input is a list of between 1 - 5000 variants (one per line) and desired commercial chip arrays to search. Input variants do not need to be on the same chromosome. 

#### Arguments
- `snps`, between 1 - 5,000 variants, using an rsID or chromosome coordinate (e.g. "chr7:24966446")
- `chip`, chip or arrays, platform code(s) for a SNP chip array, ALL_Illumina, ALL_Affy or ALL, default=ALL, use the `list_chips` utility (see below) to lookup available commercial SNP chip arrays and their codes.
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage: Multiple variants, search "ALL" available chip arrays
```{r eval=FALSE}
SNPchip(snps = c("rs3", "rs4", "rs148890987"), 
        chip = "ALL", 
        token = Sys.getenv("LDLINK_TOKEN")
       )
```

```
## WARNING: The following RS number did not have any platforms found: rs148890987, rs3.
```

```
##     RS_Number Position_GRCh37 A_SNP5.0 A_CHB2 A_250S A_SNP6.0
## 1 rs148890987  chr13:32403784        0      0      0        0
## 2         rs3  chr13:32446842        0      0      0        0
## 3         rs4  chr13:32447222        1      1      1        1
```

<br>

#### Usage: Multiple variants, search two Affymetrix arrays
```{r eval=FALSE}
SNPchip(snps = c("rs3", "rs4", "rs148890987"), 
        chip = c("A_SNP5.0", "A_CHB2"), 
        token = Sys.getenv("LDLINK_TOKEN")
       )
```

```
## WARNING: The following RS number did not have any platforms found: rs148890987, rs3.
```

```
##     RS_Number Position_GRCh37 A_SNP5.0 A_CHB2
## 1 rs148890987  chr13:32403784        0      0
## 2         rs3  chr13:32446842        0      0
## 3         rs4  chr13:32447222        1      1
```


<br>

#### Usage: Multiple variants, search all available Affymetrix arrays using, "ALL_Affy"
```{r eval=FALSE}
SNPchip(snps = c("rs3", "rs4", "rs148890987"), 
        chip = "ALL_Affy", 
        token = Sys.getenv("LDLINK_TOKEN")
       )
```

```
## WARNING: The following RS number did not have any platforms found: rs148890987, rs3.
```

```
##     RS_Number Position_GRCh37 A_SNP5.0 A_CHB2 A_250S A_SNP6.0
## 1 rs148890987  chr13:32403784        0      0      0        0
## 2         rs3  chr13:32446842        0      0      0        0
## 3         rs4  chr13:32447222        1      1      1        1
```

Output is a data frame of query variant rows (RS number), genomic coordinate (GRCh37) and genotyping chip array columns. The presence of a "1" designates the variant is present on the respective commercial genotyping array and a "0" indicates that it is not present on the genotyping array. 

***

### SNPclip

#### Function
```
SNPclip(snps, pop = "CEU", r2_threshold = "0.1", maf_threshold = "0.01", token = NULL, file = FALSE)
```
Prune a list of variants by linkage disequilibrium. Input is a list of variant RS numbers (one per line) and a population group.

#### Arguments
- `snps`, a list of between 1 - 5,000 variants, using an RS number (rsID) or chromosome coordinate (GRCh37) (e.g. "chr7:24966446"). All input variants must be on the same chromosome and match a bi-allelic variant.
- `pop`, a 1000 Genomes Project reference population, uses three letter population code, (e.g. YRI or CEU), multiple allowed, default = "CEU"
- `r2_threshold`, Used to set the R^2^ threshold for LD pruning. One of each pair of variants with a R^2^ greater than the threshold is removed. Value needs to be in the range 0 to 1. Default value is 0.1.
- `maf_threshold`, Used to set minor allele frequency (MAF) threshold for LD pruning. Variants with a MAF less than or equal to the threshold are removed. Value needs to be in the range 0 to 1. Default value is 0.01.
- `token`, LDlink provided user access token is required, default = NULL, register for a free token on the [LDlink web site.](https://ldlink.nci.nih.gov/?tab=apiaccess)
- `file`, optional character string naming a path and file for saving results. If file = FALSE, no file will be generated, default = FALSE

#### Usage:  Multiple Variants
```{r eval=FALSE}
SNPclip(snps =  c("rs3", "rs4", "rs148890987", "rs115955931"), 
        pop = "YRI", 
        r2_threshold =  "0.1", 
        maf_threshold = "0.01", 
        token = Sys.getenv("LDLINK_TOKEN")
       )
```

```
##     RS_Number       Position          Alleles
## 1         rs3 chr13:32446842 C=0.829, T=0.171
## 2         rs4 chr13:32447222 A=0.829, G=0.171
## 3 rs148890987 chr13:32403784     C=1.0, T=0.0
## 4 rs115955931 chr13:32130008 G=0.954, A=0.046
##                                             Details
## 1                                     Variant kept.
## 2 Variant in LD with rs3 (R2=1.0), variant removed.
## 3              Variant MAF is 0.0, variant removed.
## 4                                     Variant kept.
```

The output table provides details including query variant RS number, genomic position, alleles, and and details about whether the variant was kept or removed.

***

## Utilities and Examples

### list_chips

#### Function
```
list_chips()
```
Provides a data frame listing the names and abbreviation codes for available commercial SNP Chip Arrays from Illumina and Affymetrix.

#### Usage
```{r}
list_chips()
```

***

### list_pop

#### Function
```
list_pop()
```
Provides a data frame listing the available reference populations from the 1000 Genomes Project, continental or super-populations (e.g. European, African, Admixed American) and sub-populations (e.g Finnish, Gambian, Peruvian)

#### Usage
```{r}
list_pop()
```

***

### list_gtex_tissues

#### Function
```
list_gtex_tissues()
```
Provides a data frame listing the GTEx full names, `LDexpress` full names (without spaces) and acceptable abbreviation codes of the 54 non-diseased tissue sites collected for the [GTEx Portal](https://gtexportal.org/home/) and used as input for the `LDexpress` function.

#### Usage
```{r, echo=FALSE}
options(width = 100)
```

```{r}
list_gtex_tissues()
```

***

## FAQs (Frequently Asked Questions)

1. What if my access token doesn't work?
    + Please double check that the token was typed accurately.  Then, ensure the format of the function call is correct.  For example, if your alphanumeric access token is: 123abc456789, then, use it as:  
```{r eval = FALSE}
df <- LDproxy(snp = "rs456", pop = "YRI", token = "123abc456789")
```

<br>

|            If you still can not solve the problem, please email us at NCILDlinkWebAdmin@mail.nih.gov.

<br>

2. Can I set a threshold or cut-off value for R^2^ or D` values?
    + No.  *LDlinkR* functions do not include 'threshold' as an argument.  However, the returned data object can be subset using base R. For example:
   
```{r eval=FALSE}
df <- LDproxy("rs12027135", pop = "CEU",r2d = "r2", token = "YourTokenHere123")
new_df <- subset(df, R2 >= 0.8)
```

<br>

3. I need to upload hundreds of variants from a text file into LDmatrix.  Why do I get an error with the following code?
```{r eval=FALSE}
test <- read.table("variant_list.txt", header = FALSE)
LDmatrix(snps = test, pop = "CEU", r2d = "r2", token = "YourTokenHere123")
```

`Error in LDmatrix(snps = test, pop = "CEU", r2d = "r2", token = "YourTokenHere123"),  :   
Input is between 2 to 1000 variants.`

<br>

 + In the above example, 'test' is an object of class data frame and the variants are in the first column. The first column needs to be specified in order to avoid the error. The example below will yield the desired result:
   
```{r eval=FALSE}
test <- read.table("variant_list.txt", header = FALSE)
LDmatrix(snps = test[,1], pop = "CEU", r2d = "r2", token = "YourTokenHere123")
```
```{r eval=TRUE, echo=FALSE}
test <- read.table("variant_list.txt", header = FALSE)
LDmatrix(snps = test[,1], pop = "CEU", r2d = "r2", token = Sys.getenv("LDLINKR_TOKEN"))
```

<br>

4. What genome build does LDlink use for genomic coordinates?
    + All genomic coordinates are based on GRCh37/hg19.
    
<br>

5. How can I ask for help?
    + If you find that you can't answer a question or solve a problem yourself, you can email us at NCILDlinkWebAdmin@mail.nih.gov or open an issue at the *LDlinkR* GitHub repository (https://github.com/CBIIT/LDlinkR).

## Session Information
```{r}
sessionInfo()
```

